### 2016-09-18 image_setup.tex
changes

To get started, we first need to install an \ac{OS} that works properly
on all the \ac{Raspi} models. We will download and setup the image in
the testbed administrator \ac{PC} using a Debian-based distribution. We
use the popular Debian-based Raspbian Linux~\cite{raspbian} given that is
the recommended and default \ac{OS} for the \ac{Raspi}. Raspbian is made
available in two bundles: Raspbian and Raspbian Lite. The difference
between the two, is that Raspbian contains a pre-installed desktop environment
for user interaction and Raspbian Lite by default only permits to interact
through a command shell. Given that the \ac{Raspi}s in our testbed are not
connected to monitors, we decide to work with Raspbian Lite.
A desktop environment can be installed using the package manager
as it will be shown in the configuration section.

@@

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite which we have
made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are available there. The testbed
administrator has to move these files to his / her \ac{HTTP} server.
To get started, the testbed administrator must open a Linux
shell (terminal) in its \ac{PC} and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator.
%To copy the commands from this procedure to the
%testbed administrator terminal, we recommend to read this document
%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
%render properly when pasted.

@@

% Define variable]
\begin{lstlisting}[]
$ export URL="http://kom.aau.dk/project/TuneSCode/raspi/"
$ export IMAGE="2016-05-27-raspbian-jessie-lite"
$ export WORKDIR="${HOME}/Raspbian"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

% Check out the image
% The dollar hack was to fix vim syntax
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 1.3 GiB, 1387266048 bytes, 2709504 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

@@

\subsection{Image Resizing}
Given that we want to customize the root filesystem in the \ac{Raspi}s,
we need to expand the image file since 1.2 GB might not be enough to store
existing root filesystem plus additional files and software packages.
%Given that we want to customize the files stored in the \ac{Raspi}s,
%we need to resize the image file since 1.2 GB might not be enough to store
%the root filesystem with the additional packages that will be installed in this procedure.
%due to the total size of the additional packages.
Thus,
we need to increase the partition size. The following procedure illustrates
how the image and its root filesystem can be expanded by one GB.
First, to expand the image one GB, we execute:

@@

Later, we use \texttt{fdisk} with the same arguments as before to see that
the image is now one GB larger:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

@@

If the partitions commands were correct, the partition table should now
look like the following: \todot{PRINT PARTITION TABLE}

@@

\subsection{Loopback Device Setup}
After successfully resizing the image file, we use a loopback device to make
the Raspbian image available as a block device in the filesystem. For this
command to work, the tested administrator distribution must have the
\texttt{util-linux} package with version 2.21 or higher. Otherwise, the
\texttt{-P} argument of \texttt{losetup} will appear as invalid. If the
version of \texttt{losetup} can not be updated for some reason, an
alternative option for this part is presented in
Section~\ref{sec:alternative_losetup} of the Appendices.

\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
/dev/loop0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsubsection{Configuration Scripts Execution Order}
To actually make the \ac{Raspi}s change hostnames and any other considered
configurations, we have to make each \ac{Raspi} call the above scripts when
it starts up. After finishing the setup process, all the unzipped files
presented in Section~\ref{sec:configuration_files_download} should be
locally available at each \ac{Raspi} after getting the root filesystem.
We first need to run the update script before running
any other configuration scripts. To do this after boot up, we include
a call for the update script in the image \texttt{\$\{ROOTDIR\}/etc/rc.local}
file before \texttt{exit 0} using a file editor. To do this, it is required
to edit this file with root permissions. The resulting file should look like the following:

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
---------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
---------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

>>>>

To get started, we first need to install an \ac{OS} that works properly
on all the \ac{Raspi} models. We will download and setup the image in
the testbed administrator \ac{PC} using a Debian-based distribution.
\todot{We could present the two methods here (Yocto vs our method).}
We
use the popular Debian-based Raspbian Linux~\cite{raspbian} given that is
the recommended and default \ac{OS} for the \ac{Raspi}. Raspbian is made
available in two bundles: Raspbian and Raspbian Lite. The difference
between the two, is that Raspbian contains a pre-installed desktop environment
for user interaction and Raspbian Lite by default only permits to interact
through a command shell. Given that the \ac{Raspi}s in our testbed are not
connected to monitors, we decide to work with Raspbian Lite.
If required, a desktop environment can be installed using the package manager later.
%\todot{remove the rest of this line} as it will be shown in the configuration section.

@@

The latest Raspbian Lite bundle can be downloaded from the Raspbian
official webpage~\cite{raspbian}. At the time of this writing, the latest
available bundle was \texttt{2016-05-27-raspbian-jessie-lite.zip}.
To ensure that the content of the bundle does not change, this procedure
is based on that particular version of Raspbian Lite which we have
made available at~\cite{soerensen_chres_wiant_2016_154143}. All other files used in this paper are also available there. The testbed
administrator has to move these files to his / her own \ac{HTTP} server.
To get started, the testbed administrator must open a Linux
shell (terminal) in its \ac{PC} and declare the environment variables shown
in the command block below. We show the whole procedure by performing the role
of the testbed administrator.
%To copy the commands from this procedure to the
%testbed administrator terminal, we recommend to read this document
%with Adobe Acrobat Reader in Linux. Otherwise, copied characters might not
%render properly when pasted.

@@

% Define variable]
\begin{lstlisting}[]
$ export URL="http://kom.aau.dk/project/TuneSCode/raspi/"
$ export IMAGE="2016-05-27-raspbian-jessie-lite"
$ export WORKDIR="${HOME}/Raspbian"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}
\todot{change to zenodo link above}

@@

% Check out the image
% The dollar hack was to fix vim syntax
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u=sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 1.3 GiB, 1387266048 bytes, 2709504 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

@@

\subsection{Image Resizing}
Given that we want to customize the root filesystem in the \ac{Raspi}s,
we need to expand the image file since 1.2 GB might not be enough to store
the existing root filesystem plus additional files and software packages.
%Given that we want to customize the files stored in the \ac{Raspi}s,
%we need to resize the image file since 1.2 GB might not be enough to store
%the root filesystem with the additional packages that will be installed in this procedure.
%due to the total size of the additional packages.
Thus,
we need to increase the partition size. The following procedure illustrates
how the image and its root filesystem can be expanded by one GB.
First, to expand the image one GB, we execute:

@@

Later, we use \texttt{fdisk} with the same arguments as before to see that
the image is now one GB larger:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u=sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

@@

If the partitions commands were correct, the partition table should now
look like the following:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u=sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

Device                               Boot  Start     End Sectors  Size Id Type
2016-05-27-raspbian-jessie-lite.img1        8192  137215  129024   63M  c W95 FAT32 (LBA)
2016-05-27-raspbian-jessie-lite.img2      137216 4806655 4669440  2.2G 83 Linux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsection{Loopback Device Setup}
After successfully resizing the image file, we use a loopback device to make
the Raspbian image available as a block device in the filesystem. For this
command to work, the testbed administrator distribution must have the
\texttt{util-linux} package with version 2.21 or higher. Otherwise, the
\texttt{-P} argument of \texttt{losetup} will appear as invalid. If the
version of \texttt{losetup} can not be updated for some reason, an
alternative option for this part is presented in
Section~\ref{sec:alternative_losetup} of the Appendices.

\begin{lstlisting}[]
$ export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
/dev/loop0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsubsection{Configuration Scripts Execution Order}
To actually make the \ac{Raspi}s change hostnames and any other considered
configurations, we have to make each \ac{Raspi} call the above scripts when
it starts up. After finishing the setup process, all the unzipped files
presented in Section~\ref{sec:configuration_files_download} should be
locally available at each \ac{Raspi} after getting the root filesystem.
We first need to run the update script before running
any other configuration scripts. To do this after boot up, we include
a call for the update script in \texttt{\$\{ROOTDIR\}/etc/rc.local} before
\texttt{exit 0} in the file:

%before \texttt{exit 0} using a file editor. To do this, it is required
%to edit this file with root permissions. The resulting file should look like the following:

\Suppressnumber\begin{lstlisting}[]
# sed -i '/^exit 0/i bash /home/pi/rasp_config/update_rasp_config' ${ROOTDIR}/etc/rc.local
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

%\Suppressnumber\begin{lstlisting}[]
%<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
%<@\textcolor{gray}{
%---------------------------------------------}
%\Reactivatenumber @>
%...
%bash /home/pi/rasp_config/update_rasp_config
%...
%exit 0
%\end{lstlisting}
%\FloatBarrier
%\vspace{-5mm}

@@

\todot{insert sed command below instead}
\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
---------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

### 2016-09-18 overview.tex
changes

We will refer to the
testbed administrator as the person(s) in charge of setting up and
configuring the testbed with administrator privileges from the \ac{OS}
point of view. The setting and configuration procedures are
performed by the testbed administrator in a \ac{PC} running a
Linux distribution as shown in Fig.~\ref{fig:testbed_setup}. Although in principle the administrator
Linux distribution is not a restriction, we present our procedure in a
Debian-based Linux distribution. Our basic design considers to create
a customized image to store it later on a memory card for each \ac{Raspi}.
Once configured, we store the resulting image file in a \ac{HTTP} server
as backup and in case the testbed administrator requires to make new
changes to this file. In our case, we use the university \ac{HTTP} server but
any others can be used. We also put all the required
configuration files and scripts for the \ac{Raspi}s setup in the \ac{HTTP}
server so there is a single place where system setup is stored and could be
modified. This simplifies the system maintenance.
As it may not always be desirable to make persistent changes
on the \ac{Raspi}s. For example when different users are interested in
running experiments on a rebooted testbed. We later present how
to utilize stacked filesystems to enable both persistent and
temporary storage to have this capability. Its purpose is to
remove non-desired data after a reboot while keeping the original
customized image structure. This step of the procedure is optional if
the testbed administrator decides to keep only persistent changes regardless
of the testbed use. Finally, we include a set of automation, monitoring
and cross-compilation tools over the top of our system in order to simplify
the execution of repetitive and long tasks, be able to follow the progress
of long task processes and compile relevant C++ source code for the testbed
administrator.

>>>>

We will refer to the
testbed administrator as the person(s) in charge of setting up and
configuring the testbed with administrator privileges from the \ac{OS}
point of view. The setting and configuration procedures are
performed by the testbed administrator in a \ac{PC} running a
Linux distribution as shown in Fig.~\ref{fig:testbed_setup}. Although in principle the administrator
Linux distribution is not a restriction, we present our procedure in a
Debian-based Linux distribution. Our basic design considers to create
a customized image to store it later on a memory card for each \ac{Raspi}.
Once configured, we store the resulting image file in a \ac{HTTP} server
as backup and in case the testbed administrator requires to make new
changes to this file.
%In our case, we use the university \ac{HTTP} server but any others can be used.
In our case, we store all files at Zenodo~\cite{soerensen_chres_wiant_2016_154143}, but the testbed administrator should copy the our files to his / her own \ac{HTTP} server to get read/write permissions.
We also put all the required
configuration files and scripts for the \ac{Raspi}s setup in the \ac{HTTP}
server so there is a single place where system setup is stored and could be
modified. This simplifies the system maintenance, as it may not always be desirable to make persistent changes
on the \ac{Raspi}s. For example when different users are interested in
running experiments on a rebooted testbed. We later present how
to utilize stacked filesystems to enable both persistent and
temporary storage to have this capability. Its purpose is to
remove non-desired data after a reboot while keeping the original
customized image structure. This step of the procedure is optional if
the testbed administrator decides to keep only persistent changes regardless
of the testbed use. Finally, we include a set of automation, monitoring
and cross-compilation tools over the top of our system in order to simplify
the execution of repetitive and long tasks, be able to follow the progress
of long task processes and compile relevant C++ source code for the testbed
administrator.
